report_dendro

### 2017-12-20 reports/report_dendro.Rmd
add report dendro

---
title: "material and methods"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: ../references.bib
templat
csl: ../geb.csl
---

@@

# Materials and methods

## Species and study site

The Pyrenean oak (*Quercus pyrenaica* Willd.) forests extend through south-western France and the Iberian Peninsula [@Franco1990] reaching its southern limit in north of Morocco. In the Iberian Peninsula these forests live under meso-supramediterranean and mesotemperate areas and subhumid, humid and hyperhumid ombroclimate [@RivasMartinez2002] living on siliceous soils, or soils poor in basic ions [@Vilches2014]. *Q. pyrenaica* requires between 650 and 1200 mm of annual precipitation and a summer minimal precipitation between 100 and 200 mm (Martínez-Parras and Molero-Mesa 1982, García and Jiménez 2009), with summer rainfall being a key factor in the distribution of the species (Gavilán et al. 2007, Río et al. 2007).

This species reaches its southernmost European limit at Sierra Nevada, a high-mountain range located in southern Spain (37°N, 3°W) with elevations of between 860 m and 3482 m *a.s.l.*. The climate is Mediterranean, characterized by cold winters and hot summers, with pronounced summer drought (July-August). There are eight oak patches (2400 Has) identified (**Figure 1**) in this mountain range, ranging between 1100 and 2000 m *a.s.l.* and generally associated to major river valleys. Sierra Nevada is considered a glacial refugia for deciduous *Quercus* species during glaciation (Brewer et al. 2002, Olalde et al. 2002, Rodríguez-Sánchez et al. 2010) and these populations are considered as a rear edge of the habitat distribution, which is important in determining habitat responses to expected climate change (Hampe and Petit 2005).

:red_circle: `duda aqui` Varias referencias hablan de los años 2005 y 2012 como  extremadamente secos. Pero habría que hacer alguna referencia y/o análisis. Tengo dudas de si hemos de analizar (e incluir) que efectivamente los años 2005 y 2012 fueron caracterizados por un extrema sequía, por lo que habría que  incluyendo referencia a apéndice

* O quizá un apartado llamado Drought episodes (similar a esto https://www.nature.com/articles/srep28269)

The populations of Pyrenean oak forests at Sierra Nevada are considered relict forests (Melendo and Valle 2000, Vivero et al. 2000), undergoing intensive anthropic use in the last few decades (Camacho-Olmedo et al. 2002, Valbuena-Carabaña et al. 2010). In fact, the status of conservation of this species for southern Spain is "Vulnerable" (Vivero et al. 2000). The relict presence of this species in Sierra Nevada is related both to its genetic resilience as well as to its high intraspecific genetic diversity (Valbuena-Carabaña and Gil 2013). However, they are also expected to suffer the impact of climate change, due to their climate requirements (wet summers). Thus, simulations of the climate change effects on this habitat forecast a reduction in suitable habitats for Sierra Nevada (Benito et al. 2011).

## Datos de sequía.

* :red_circle: Meter aquí algunos datos de sequia, similar a lo planteado por Gazol

### Greenness data

To characterize the vegetation greenness of *Q. pyrenaica* we used the *Enhanced Vegetation Index* (EVI) derived from MOD13Q1 product obtained by the *Moderate Resolution Imaging Spectroradiometer* (MODIS) sensor [@Didan2015]. EVI and NDVI (*Normalized Difference Vegetation Index*) are the most common greenness vegetation indices. We used EVI instead of NDVI because EVI is more sensitive to changes in high-biomass areas (a serious shortcoming of NDVI); EVI reduces the influence of atmospheric conditions on vegetation index values, and EVI corrects for canopy background signals [@Huete2002; @Krapivin2015; @Cabello2012].

EVI data consits of 16-day maximun value composite images (23 per year) of the EVI value with a spatial resolution of 250 m x 250 m. We selected the pixels covering the distribution of *Q. pyrenaica* forests in Sierra Nevada (*n* = 928 pixels). MODIS EVI Data from Collection 6 were obtanied using Google Earth Engine platform for the period 2000 - 2016. A data filtering was applied to select EVI valid values. The filtering was done using quality flags and VI Usefulness Indices accompanying the EVI data. We filter out those values affected by high content of aerosols, clouds, shadows, snow or water.

Each 1 × 1 km2 16-day composite EVI value is considered valid when (a) EVI data is produced—'MODLAND_QA' equals 0 (good quality) or 1 (check other QA), (b) VI usefulness is between 0 and 11, (c) clouds are absent—'adjacent cloud detected' (0), 'mixed clouds' (0) and 'possible shadow' (0), and (d) aerosol content is low or average—'aerosol quantity' (1 or 2). Note that 'MODLAND_QA' checks whether EVI is produced or not, and if produced, its quality is good or whether other quality flags should also be checked. Besides, VI usefulness indices between 0 to 11 essentially include all EVI data. Thus, these two conditions serve as additional checks.

and then a quality assessment was carried out to filter the ... [@Reyes2015]

    * According to @Reyes2015 we must consider the shadow in the mountain, but we can discard the filter of adjacent clouds. On the other hand, the use of EVI mean is highly stable under the use of any filter [@Reyes2015]

The presence of clouds (adjacent clouds, mixed clouds and shadows) 'obscures' the surface in a radiometric sense, thus corrupting inferred EVI values. In addition, two types of aerosol loadings typically corrupt EVI—climatology and high aerosols. Use of aerosol climatology

The EVI data are geometrically and atmospherically corrected and include information about the quality ass....

`$NOTA$`: NDVI sirve para estimar la producción primaria neta. Existen diferentes estudio que han evaluado el efecto de la sequía sobre la producción primaria neta utilizando NDVI.

These data are geometrically and atmospherically corrected, and include an index of data quality (reliability, which range from 0 – good quality data – to 4 – raw data or absent for different reasons) based on the environmental conditions in which the data was recorded

We first used the Quality Assesment (QA band) information of this product to filter out those values affected by high content of aerosols, clouds, shadows, snow or water; and then a quality assessment was carried out to filter the ... [@Reyes2015]

:red_circle: reescribir esto de la calidad. Ver Samanta et al 2012 y como describe el proceso de calidad

OJITO---------- reescribir esto de arriba

After the filter out process, we built the annual EVI profile for each pixel and then computed the EVI's annual mean values and the EVI anomaly for each pixel for the period 2000 - 2015. (:red_circle: Hemos seleccionado EVI medio, además de por los consejos que me ha dicho Domingo, porque he comprobado que existe una correlación entre el evi medio y el evi estacional, sobre todo el de verano. Ver esto: [https://github.com/ajpelu/qpyr_modis_resilience/blob/master/analysis/prepare_modis_qa.md](https://github.com/ajpelu/qpyr_modis_resilience/blob/master/analysis/prepare_modis_qa.md). Además presenta alta correlaciones significativas con el EVI de verano: 0.88; de primavera: 0.76 y anual: 0.81)

> Procedimiento de Filtrado de datos (ver [https://github.com/ajpelu/qpyr_modis_resilience/blob/master/analysis/prepare_modis_qa.md](https://github.com/ajpelu/qpyr_modis_resilience/blob/master/analysis/prepare_modis_qa.md))

* Información contenida en banda QA.

    * Nos quedamos con pixeles marcados como Good Data (57.89 %)
    * Filter out los marcados como Snow/Ice y/o Cloudy (2.57 + 7.08 = 9.65 %)
    * Pixeles marcados como Marginal Data (32.33 %) (ver siguiente paso)

* Explorar distribución temporal y analizar banda QA Detailed y llevar a cabo un filtrado siguiendo las especificaciones de @Reyes2015.

    * Vemos los composites marcados con Aerosoles, Adjacent cluods, y Shadow.  
    * According to @Reyes2015 we must consider the shadow in the mountain, but we can discard the filter of adjacent clouds. On the other hand, the use of EVI mean is highly stable under the use of any filter [@Reyes2015]

* Finalmente nos hemos quedado con las siguientes cifras. De un total de 360064 images composites for the study zone were downloaded (928 x 20 x 1 + 928 x 23 x 16 = 360064), tras el filtrado, nos quedamos con 286825 (79.65 %)

To explore the effect of drought events on greenness we calculated the EVI standardized anomaly ($EVI~sa$) pixel-by-pixel, since it minimizes biases in the evaluation of anomalies, providing more information about the magnitude of the anomalies [@Samanta2012, @Gao2016]. For each pixel we averaged all the EVI valid values within a year (:red_filter: see quality filter), and then the standardized anomaly was computed as:

$$\mathrm{EVI_{sa,\mathit{i}}}= \frac{\mathrm{EVI_{mean,\mathit{i}}-EVI_{mean,ref}}}{\sigma_{\mathrm{ref}}}$$
where $\mathrm{EVI_{sa,\mathit{i}}}$ is the EVI standardized anomaly for the year $i$;  $\mathrm{EVI_{mean,\mathit{i}}}$ the annual mean value of EVI for the year $i$; $\mathrm{EVI_{mean,ref}}$ the average of the annual EVI values for the period of reference (all except $i$ year), and $\sigma_{\mathrm{ref}}$ the standard deviation for the reference period.

Each pixel was categorized according the EVI standardized anomalies as "greening" (EVI standardized anomalies greater than + 1), "browning" (EVI standardized anomalies less than - 1) or "no changes" (EVI standardized anomalies between − 1 and  + 1)

@@

### Tree sampling 
Samplig was carried during autumn of 2016. Trees were sampled at two locations in contrasting slopes of Sierra Nevada: San Juan (SJ; northern site) and Cáñar (CA; southern site) (Figure 1; Table 1). Two elevations were sampled for the southern site (CA-Low and CA-High). All the sites were oak monospecific and representatives of two of the three the population's cluster identified for the specie in this mountain range (:red_circle: mejorar; citar Pérz-Luque et al..). In each site between 15 and 20 dominant trees were randomly selected. Two cores of 5 mm of diameter were taken from each at breast heith (1.3 m) using an increment borer. For each tree, diameter at breat height (DBH) and total height were recorded. Increment cores were air dried, glued onto wooden mounts and sanded. Annual radial growth (ring width, RW) were measured with a LINTAB measuring device (Rinntech, Heidelberg, Germany) coupled to a stereomicroscope, with an accuracy of 0.01 mm. Individual ring series were visually and statistically cross-dated with TSAP software (Rinntech, Heidelberg, Germany), using the statistics Gleichläufigkeit (GLK), t-value and the crossdating index (CDI). Validation of the croos-dating was done using COFECHA software [@Holmes1983].

## Dendrochronological methods
* incluir citas a Frits; Cooks ... 
* site cronologies
* BAI 
* GC Nowacki??

@@

>>>>

---
title: "report_dendro"
output:  
  md_document:
    toc: true
    variant: markdown_github
bibliography: ../references.bib
templat
csl: ../geb.csl
---

@@

```{r setup, echo=FALSE}
show_text <- FALSE
```

# Cronologías

* Para cada sitio construimos una cronología utilizando las series de BAI. 
* Hemos utilizado la media aritmética

## Cronologías San Juan (northern face)

![]()

@@

### Tree sampling 
Samplig was carried during autumn of 2016. Trees were sampled at two locations in contrasting slopes of Sierra Nevada: San Juan (SJ; northern site) and Cáñar (CA; southern site) (Figure 1; Table 1). Two elevations were sampled for the southern site (CA-Low and CA-High). All the sites were oak monospecific and representatives of two of the three the population's cluster identified for the specie in this mountain range (:red_circle: mejorar; citar Pérz-Luque et al..). In each site between 15 and 20 dominant trees were randomly selected. Two cores of 5 mm of diameter were taken from each at breast heith (1.3 m) using an increment borer. For each tree, diameter at breat height (DBH) and total height were recorded.

### Dendroecological analyses 
Tree cores were air dried, glued onto wooden mounts and sanded. Annual radial growth (ring width, RW) were measured with a LINTAB measuring device (Rinntech, Heidelberg, Germany) coupled to a stereomicroscope, with an accuracy of 0.01 mm. Individual ring series were visually and statistically cross-dated with TSAP software (Rinntech, Heidelberg, Germany), using the statistics Gleichläufigkeit (GLK), t-value and the crossdating index (CDI). Validation of the croos-dating was done using COFECHA software [@Holmes1983].

The growth trends were analyzed at different time scales. The long-term growth response was analyzed using basal area increment (hereafter BAI, $\mathrm{cm^2 \cdot year^{-1}}$). BAI represents a more accurate indicator of growth than ring-width, since it removes variation in growth attributatle to increasing stem circumference [@Biondi2008]. We used measured dbh and raw nondetrened ring-widths to compute BAI by subtracting twice the annual ring width from the annual diameter, starting from the measured diameter outside the bark [@Piovesan2008]. We calculated a mean BAI series for each individual tree. 
To study the rsponse of growth to the inter-annual variability of climate (short-term response) we used pre-withened residual chronologies (growth index, GIres) calculated from biweigth means of ratios between raw growth measurements and individual cubic splines with a 50 % frecuency cutoff at 30 years [@Fritts1976]

```{r echo=FALSE, results='asis', eval=show_text}
cat("
 * BAI represents a more accurate indicator of growth than ring-width, since it reduces the variation caused by adding volume to a circular stem 
 * BAI is a more meaningful indicator of growth than ring width because it removes variation in growth attributale to increasing circumference. 
 * BAI is more closely related to biomass increment 
 * BAI removes variation in growth attributatle to increasing stem circumference and captures changes in growth better than linear measures as tree-ring width [@Biondi2008] 
")
```

@@

* incluir citas a Frits; Cooks ...

* BAI 
* site cronologies
* GC Nowacki??
